Response Letter for Manuscript ID BIOM2020344P - 'Efficient Detection and Classification of Epigenomic Changes Under Multiple Conditions

### 2021-02-24 Latex/letter/letter.Rmd
Update letter.Rmd
typo fixes

> We believe that Spearman correlation and LFC are not good metrics to compare peak callers with broad marks for two reasons. First, methods that tend to call narrow peaks within brod regions of enrichment will exhibit an unrealistic advantage in performance in terms of LFC. This is so because these methods tend to call as narrow peaks any genomic window exhibiting excessively high read count enrichment (or, LFC in differential enrichment). For broad marks, interest lies not on these single punctate windows but instead in broad regions of enrichment. Second, and due to a similar reason, these methods will also exhibit unrealistic advantage in performance in terms of Spearman correlation because broad regions of enrichment are much noisier than punctate regions and, thefore, read count correlations calculated from broad peaks will be attenuated in comparison to correlations calculated from narrow and punctate windows.

> Regarding the use of 'matched differentially expressed genes as the gold standard to evaluate the differential transcribed genes called from differential H3K36me3 activity', we did not observe a sufficient correspondance between gene expression levels and H3K36me3 enrichment for the cancer cell lines analyzed in our paper. For instance, we have found several instances were genes were expressed but no enrichment of H3K36me3 was observed. Below, we present some of these problematic cases obtained from the UCSC Genome Browser. Note in Figure 1, for example, that both cell lines Helas3 and Hepg2 were not enrichment for H3K36me3 in the rightmost region of the plot while the associated gene body was clearly expressed. Note in Figure 2, for example, the lack of agreement between the transcription level of the reported gene and the enrichment for H3K36me3 in cell line Hepg2. Due to these reasons, we decided not to use gene expression levels when benchmarking H3K36me3 as suggested by the reviewer.

@@

> We believe our method is quite comparable to other DPC in short differential peak calling. These results are shown in Figure 4 of our paper as well as in the Web Appendix (Web Figures 25-27). In terms of LFC and Spearman correlation, our method performed quite comparably to ChIPComp, a widely used method for narrow marks. Examples in Figure 4B, 4D, and 4F clearly shows that our model is correctly calling differential peaks similar to other methods.. We did see, however, a slight increase in memory consumption of 5GB when analyzing such marks due to the higher resolutions of 250bp. However, such an increase is still well within the available resourses of most high performance computing clusters and even in some personal laptops nowadays.

@@

> We thank the reviewer for the constructive comment. As we replied to the first anonymous reviewe, we did not observe a sufficient agreement between gene expression levels for the analyzed cell lines in our paper (see Figures 1 and 2 above). In fact, not all the analyzed marks in this paper are associated with the expression level of the neared gene. For instance, H3K27me3 and EZH2 are two epigenomic marks known to exhibit read enrichment in intergenic regions without any nearby gene body. We believe that this analysis would be suboptimal to the one presented in this paper.

>>>>

> We believe that Spearman correlation and LFC are not good metrics to compare peak callers with broad marks for two reasons. First, methods that tend to call narrow peaks within broad regions of enrichment will exhibit an unrealistic advantage in performance in terms of LFC. This is so because these methods tend to call as narrow peaks any genomic window exhibiting excessively high read count enrichment (or, LFC in differential enrichment). For broad marks, interest lies not on these single punctate windows but instead in broad regions of enrichment. Second, and due to a similar reason, these methods will also exhibit unrealistic advantage in performance in terms of Spearman correlation because broad regions of enrichment are much noisier than punctate regions and, thefore, read count correlations calculated from broad peaks will be attenuated in comparison to correlations calculated from narrow and punctate windows.

> Regarding the use of 'matched differentially expressed genes as the gold standard to evaluate the differential transcribed genes called from differential H3K36me3 activity', we did not observe a sufficient correspondance between gene expression levels and H3K36me3 enrichment for the cancer cell lines analyzed in our paper. For instance, we have found several instances were genes were expressed but no enrichment of H3K36me3 was observed. Below, we present some of these problematic cases obtained from the UCSC Genome Browser. Note in Figure 1, for example, that both cell lines Helas3 and Hepg2 did not show enrichment for H3K36me3 in the rightmost region of the plot while the associated gene body was clearly expressed. Note in Figure 2, for example, the lack of agreement between the transcription level of the reported gene and the enrichment for H3K36me3 in cell line Hepg2. Due to these reasons, we decided not to use gene expression levels when benchmarking H3K36me3 as suggested by the reviewer.

@@

> We believe our method is quite comparable to other DPC methods in short differential peak calling. These results are shown in Figure 4 of our paper as well as in the Web Appendix (Web Figures 25-27). In terms of LFC and Spearman correlation, our method performed quite comparably to ChIPComp, a widely used method for narrow marks. Examples in Figure 4B, 4D, and 4F clearly shows that our model is correctly calling differential peaks similar to other methods. We did see, however, a slight increase in memory consumption of 5GB when analyzing such marks due to the higher resolutions of 250bp. However, such an increase is still well within the available resourses of most high performance computing clusters and even in some personal laptops nowadays.

@@

> We thank the reviewer for the constructive comment. As we replied to the first anonymous reviewer, we did not observe sufficient agreement between gene expression levels for the analyzed cell lines in our paper (see Figures 1 and 2 above). In fact, not all the analyzed marks in this paper are associated with the expression level of the neared gene. For instance, H3K27me3 and EZH2 are two epigenomic marks known to exhibit read enrichment in intergenic regions without any nearby gene body. We believe that this analysis would be suboptimal to the one presented in this paper.
